% Punctuation marks
\def \Pa #1{(#1)}
\def \Cl {:}
\def \Cm {,}
\def \Ec {,}
\def \Pr {.}
\def \Pt {.}
\def \Md {---}
\def \Nd {--}
\def \Sc {;}
\def \Sp { }
\def \Ts {,}
\def \Ws {}
% Words and word combinations
\def \Iso {en}
\def \Tsp #1{Senior #1 Samples of the\\International Linguistics Olympiad, China}
\def \Hsp #1{Senior #1 Samples of the International Linguistics Olympiad, China}
\def \Tct #1#2{%
  \ifpast#1 of the\\#2 International Linguistics Olympiad, China%
  \else Senior #1 of the\\#2 International Linguistics Olympiad, China\fi}
\def \Hct #1#2{%
  \ifpast#1 of the #2 International Linguistics Olympiad, China%
  \else Senior #1 of the #2 International Linguistics Olympiad, China\fi}
\def \Mdy #1#2#3{#1 #2, #3}
\def \Opr {Open Round}
\def \Ivr {Invitational Round}
\def \Idc {Individual Contest}
\def \Tmc {Team Contest}
\def \Pbl {Problem}
\def \Pbs {Problems}
\def \Sol {Solution}
\def \Sls {Solutions}
\def \Ans {Answer Sheet}
\def \Doc #1#2{#1 #2}
\def \Pri #1{Problem #1}
\def \Pts #1{#1 points}
\def \Atc #1{a #1}
\def \Atv #1{an #1}
\def \Nos {numbers}
\def \Aeq {an arithmetic equality}
\def \Aes {arithmetic equalities}
\def \Eqs {equalities}
\def \Wds {words}
\def \Wcs {word combinations}
\def \Sfs {sentence fragments}
\def \Stc {sentence}
\def \Sts {sentences}
\def \Inl #1#2{#1 in #2}
\def \Ptl #1{partially #1}
\def \Wri #1#2{#1 written #2}
\def \Wro #1#2{#1 #2}
\def \Wru #1#2{#1 written #2}
\def \Ins #1{in #1}
\def \Inn {in numerals}
\def \Usg #1{using #1}
\def \And #1#2{#1 and #2}
\def \Bnd #1#2{#1 and #2}
\def \Cnd #1#2{#1 and #2}
\def \Que #1#2{\And{#1}{#2}}
\def \Atq #1{, and #1}
\def \Iao #1{#1 in arbitrary order}
\def \Sua #1{their #1}
\def \Mnt #1{most natural #1}
\def \Psb #1{possible #1}
\def \Txs {phonetic transcriptions}
\def \Trs #1{#1 translations}
\def \Nog #1{no greater than #1}
\def \Lss #1{less than #1}
\def \Grt #1{greater than #1}
\def \Nol #1{no less than #1}
\def \Ths #1{these #1}
\def \One #1{one of #1}
\def \Etc {, etc}
\def \Nsl {nasalization}
\def \Slb {syllable break}
\def \Eds {Editors}
\def \Txt {English text}
\def \Eic {editor-in-chief}
\def \Ted {technical editor}
\def \Nam {Name}
\def \Stn {Seat \#}
\def \Pbn {Problem \#}
\def \Pgn {Page \#}
\def \Dne {DNE}
\def \Otw {otherwise}
\def \Bef #1{before a #1}
\def \Aft #1{after a #1}
\def \Mod #1#2{#1 #2}
\def \Snd {sound}
\def \Csn {consonant}
\def \Vow {vowel}
\def \Uvc {voiceless}
\def \Vce {voiced}
\def \Plo {plosive}
\def \Nas {nasal}
\def \Stm {stem}
\def \Vrb {v.}
\def \Nun {n.}
\def \Adj {adj.}
\def \Itr {intr.}
\def \Tra {tr.}
\def \Sng {singular}
\def \Plr {plural}
\def \Nps {npst.}
\def \Pst {pst.}
\def \Prs {prs.}
\def \Nfu {nfut.}
\def \Fut {fut.}
\def \Dft {default}
\def \Prg {prog.}
\def \Ipf {ipfv.}
\def \Pfv {pfv.}
\def \Nom {nominative}
\def \Acc {accusative}
\def \Gen {genitive}
\def \Obl {oblique}
\def \Cmp {comparative}
\def \Loc {locative}
\def \Fsg {I}
\def \Ssg {you (sg.)}
\def \Tsg {he}
\def \Fpl {we}
\def \Spl {you (pl.)}
\def \Tpl {they}
% Sentences
\def \Rule {You should describe any patterns or rules that you identified in the data unless the problem explicitly instructs you otherwise.}
\def \Gbis #1{Here is #1:}
\def \Gbre #1{Here are some #1:}
\def \Giao #1{Here are some #1 in arbitrary order:}
\def \Dtma {Determine the correct correspondences between them.}
\def \Dtmb {Determine the correct correspondences:}
\def \Fila {Fill in the gaps among them.}
\def \Filb {Fill in the gaps:}
\def \Trla #1#2{Translate #1 into #2.}
\def \Trlb #1{Translate into #1:}
\def \Tnwa #1#2{Translate #1 into #2 in the most natural way.}
\def \Tnwb #1{Translate into #1 in the most natural way:}
\def \Tpwa #1#2{Translate #1 into #2 in all possible ways.}
\def \Tpwb #1{Translate into #1 in all possible ways:}
\def \Wria #1#2{Write #1 in #2.}
\def \Wrib #1{Write in #1:}
\def \Wrna #1{Write #1 in numerals.}
\def \Wrnb {Write in numerals:}
\def \Wroa #1#2{Write #1 out in #2.}
\def \Wrob #1{Write out in #1:}
\def \Wrua #1#2{Write #1 using #2.}
\def \Wrub #1{Write using #1:}
\def \Expa {Explain your solution.}
\def \Alle #1{All numbers in these equalities are #1.}
\def \Allp #1{All numbers in this problem are #1.}
\def \Coin #1#2{The answer to #1 should coincide with #2.}
\def \Noex {No additional explanation besides the answers is required, nor will be marked.}
\def \Isli #1{#1 is a language isolate.}
\def \Infm #1#2{#1 belongs to #2.}
\def \Inbr #1#2#3{#1 belongs to #2 of #3.}
\def \Ingr #1#2#3#4{#1 belongs to #2 of #3 of #4.}
\def \Hist #1#2{It was spoken in #1 around \mbox{#2} years ago.}
\def \Exti #1#2{It was spoken by the #1 ethnic groups in #2.}
\def \Glob #1{It is spoken by approx. \mbox{#1} people all over the world.}
\def \Stat #1#2{It is spoken by approx. \mbox{#1} people in #2.}
\def \Ethn #1#2#3{It is spoken by approx. \mbox{#1} people of the #2 ethnic groups in #3.}
\def \Isac #1{#1 is a consonant.}
\def \Arec #1{#1 are consonants.}
\def \Isav #1{#1 is a vowel.}
\def \Arev #1{#1 are vowels.}
\def \Indi #1#2{The mark #1 indicates #2.}
\def \Long #1{The mark #1 indicates that the preceding vowel is long.}
\def \Dvdl {The doubling of a vowel denotes length.}
\def \Anop #1{#1 are names of people.}
\def \Gdlk {Good luck!}
% Languages and scripts
\def \TXT {\ENG}
\def \ENG {English}
\def \ZHO {Chinese}
\def \LATN {the Latin script}
\def \ARABN {Arabic numerals}
% Families
\def \AAV {the Austoasiatic family}
\def \AFA {the Afro-Asiatic family}
\def \ALV {the Atlantic-Congo family}
\def \INE {the Indo-European family}
\def \MAP {the Austronesian family}
\def \SIT {the Sino-Tibetan family}
\def \TAI {the Kra-Dai family}
\def \TUW {the Tungusic family}
\def \XND {the Na-Dene family}
% Countries
\def \AU {Australia}
\def \BD {Bangladesh}
\def \BO {Bolivia}
\def \BR {Brazil}
\def \BT {Bhutan}
\def \CM {Cameroon}
\def \CN {China}
\def \DE {Germany}
\def \FJ {Fiji}
\def \GB {the UK}
\def \GE {Georgia}
\def \HU {Hungary}
\def \ID {Indonesia}
\def \IL {Israel}
\def \IN {India}
\def \KG {Kyrgyzstan}
\def \KH {Cambodia}
\def \LK {Sri Lanka}
\def \MM {Myanmar}
\def \MN {Mongolia}
\def \MX {Mexico}
\def \NG {Nigeria}
\def \PE {Peru}
\def \PG {Papua New Guinea}
\def \PH {the Philippines}
\def \RU {Russia}
\def \RO {Romania}
\def \SK {Slovakia}
\def \TH {Thailand}
\def \TJ {Tajikistan}
\def \TZ {Tanzania}
\def \TR {Turkey}
\def \US {the USA}
\def \UZ {Uzbekistan}
\def \VN {Vietnam}
% People
\def \HhLi {\mbox{Huihan Li}}
\def \YeLq {\mbox{Ye Liu}}
\def \YyLq {\mbox{Yuyang Liu}}
\def \YmLo {\mbox{Yiming Luo}}
\def \OkQq {\mbox{Ao Qiu}}
\def \YpSp {\mbox{Yupei Sun}}
\def \ZjWh {\mbox{Ziji Wang}}
\def \JnXu {\mbox{Jin Xu}}
\def \LnYe {\mbox{Lin Ye}}
\def \SyYu {\mbox{Shuyue Yu}}
\def \LhVg {\mbox{Linghui Zheng}}
